An algorithm for a spatial mapping using a hexagon tilegram, with application to Australia disease maps

### 2019-01-30 algorithm_radix.Rmd
Don't evaluate read_shape function call

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

@@

**The following must be provided to `create_hexmap`:**
```{r table_necessary}
knitr::kable(tibble::tibble(parameter = c('shp', 
  'shp_path',
  'sf_id',
  'focal_points'),
  description = c('a shape file', 'character string location of shape file', 'name of a unique column that distinguishes areas',
    'a data frame of reference locations when allocating hexagons, capital cities of Australia are provided as capital_cities'
    )))

@@

```{r data}
library(tidyverse)
library(sugaRbag)

shp_path <- "data/lga_2011.rda"
sf_id <- "LGA_CODE11"
focal_points <- sugaRbag::capital_cities
```

@@

```{r}
width <- 15 # arbitrary angle width
```

@@

- To utilise a shape file from a directory on your device, provide the file path for .shp file.
- If the shape file is an .Rda file, it is loaded into the current environment.

@@

We will use the Local Government Areas of Hobart as an example:

```{r}
# Read in shape information
nsw_shp <- read_shape(shp_path = shp_path)

# Convert to a simple features object
nsw_lga <- sf::st_as_sf(nsw_shp)

# Select only desired LGA areas, those from nswmania.
nsw_lga <- nsw_shp %>% filter(STE_NAME11 == "New South Wales")
nsw_lga[[sf_id]] <- droplevels(as.factor(nsw_lga[[sf_id]]))

## Not necessary when using `create_hexmap`:
## Derive projection information
crs_info <- sf::st_crs(nsw_lga)
proj4string <- crs_info$proj4string
epsg <- crs_info$epsg

@@

nsw_lga <- nsw_lga %>%
    sf::st_transform(., crs = projstring)

## Use the create_centroids function to find polygon centroids:
centroids <- create_centroids(shp_sf = nsw_lga, sf_id = sf_id, projstring = projstring, verbose = FALSE)

@@

```{r, fig.cap="nswmania Local Government Areas and derived centroids."}
fort_nsw <- fortify_sfc(nsw_lga, projstring = projstring)
ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude), data= centroids, colour = "#b2df8a") + theme_void() + coord_equal()
```

@@

```{r}
## Not necessary when using `create_hexmap`:
## create the bounding box from the set of polygon centroids:
bbox <- tibble::tibble(min = c(min(centroids$longitude), min(centroids$latitude)),
        max = c(max(centroids$longitude), max(centroids$latitude)))

@@

```{r}
grid <- tibble::as_tibble(expand.grid(hex_long = seq(bbox$min[1] - buffer_dist,
    bbox$max[1] + buffer_dist,
    hex_size),
    hex_lat = seq(bbox$min[2] - buffer_dist,
        bbox$max[2] + buffer_dist,
        hex_size)))
```

```{r}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude), data= centroids, colour = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75) 
```

@@

```{r}
# Find every second latitude
shift_lat <- grid %>% dplyr::select(hex_lat) %>%
    dplyr::distinct() %>%
    dplyr::filter(dplyr::row_number() %% 2 == 1) %>% unlist()

@@

```{r}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroids, fill = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75) 
```

@@

```{r}
full_grid <- grid %>%
    mutate(hex_long_int = dense_rank(hex_long)-1,
        hex_lat_int = dense_rank(hex_lat)-1)
```

@@

```{r}
grid <- create_buffer(centroids, full_grid, hex_size, buffer_dist)
```

@@

### Step 3: Filtering the grid
Only the grid points between the rolling average of the minimum and maximum centroid values are kept, for each row and column of the grid.

```{r}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroids, fill = "#b2df8a", fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = full_grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75)
```

@@

```{r}
# Split the centroid data set
centroids <- centroids %>% 
  nest(longitude, latitude) %>% 
  mutate(closest = map(data, closest_focal_point, focal_points = focal_points)) %>% 
  unnest(data, closest) %>% 
  arrange(focal_distance)
```

@@

```{r}
## Not necessary when using `create_hexmap`:
## Consider hexagon points up to a certain distance from centroid

@@

```{r, eval = FALSE}
hexmap_allocation <- allocate(centroids = centroids,
        hex_grid = grid,
        hex_size = hex_size,
        filter_dist = filter_dist,
        width = width,
        focal_points = focal_points,
        verbose = TRUE,
        id = sf_id)
```

@@

```{r}
s_centroids <- centroids %>% arrange(focal_distance)

@@

```{r}
# keep value to reset expanded distances
expand_dist <- filter_dist
```

@@

```{r}
# filter for only the available hex grid points
hex_grid <- grid %>% filter(!assigned)
```

### Step 2: Filter the grid points for those closest to the centroid
```{r}
# filter grid for avaiable points
centroid1 <- centroids %>% head(1)

@@

Create a box filter
```{r, echo = FALSE}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroid1, fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)

@@

Create circle filter
```{r}
    hex_grid <- hex_grid %>%
        rowwise %>%
        mutate(
        hex_lat_c = hex_lat - flat,
        hex_long_c = hex_long - flong) %>%
        mutate(hyp = ((hex_lat_c^2) + (hex_long_c^2))^(1/2))

@@

Filter for angle within circle, relates hexagon location to focal point, Hobart.

@@

The allocation is returned and combined with the data relating to each polygon.
```{r}
hex_points_df <- centroid_allocation %>% 
  group_by(LGA_CODE11) %>%
  nest() %>%
  mutate(hex_points = 
      map(.x = data,.f = function(a) { 
        fortify_hexagon(data = a, hex_size = hex_size)})) %>%
  unnest(data) %>%
  unnest(hex_points)
```

@@

```{r}
ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") +
  theme_void() + coord_equal() +
    geom_hex(aes(x = hex_long, y = hex_lat, fill = AREA_SQKM), data = hexmap_lga, 
      position = "identity", stat = "identity")
```

@@

Examples of Melbourne, nswmania, Australia
At different levels, distance from original centroids, change in area

LGA of all Australia
SA2 for nswmania

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = FALSE)
```

@@

**The following must be provided to `create_hexmap`:**
```{r table_necessary}
knitr::kable(tibble::tibble(parameter = c('shp', 
  'shp_path',
  'sf_id',
  'focal_points'),
  description = c('an Rdata object containing the polygon information', 'character string location of the shape file that contains the polygon information', 'name of a unique column that distinguishes areas',
    'a data frame of reference locations when allocating hexagons, capital cities of Australia are provided as capital_cities'
    )))

@@

```{r libraries}
library(tidyverse)
# To install:
# devtools::install_github("srkobakian/sugaRbag")
library(sugaRbag)
```

The polygon set of Local Government Areas in Australia in 2011 is stored in the sugaRbag package as `lga_2011`.

```{r data}
sf_id <- "LGA_CODE11"
focal_points <- sugaRbag::capital_cities
```

@@

@@

- If the shape file is an .Rda file, it is loaded into the current environment.
- To utilise a custom shape file from a directory on your device, provide the file path for .shp file.

```{r, eval = FALSE}
# Read in shape information
shp <- read_shape(shp_path = shp_path)

# Convert to a simple features object
shp_sf <- sf::st_as_sf(shp)
```

@@

We will use the Local Government Areas of Australia in the `lga_2011` data set, and take only those from New South Wales as an example:

```{r}
shp_sf <- sugaRbag::lga_2011

nsw_lga <- shp_sf %>% dplyr::filter(STE_NAME11 == "New South Wales")
# Select only desired LGA areas, those from nsw.
nsw_lga[[sf_id]] <- droplevels(as.factor(nsw_lga[[sf_id]]))
```

```{r project}
## Not necessary when using `create_hexmap`:
## Derive projection information
crs_info <- sf::st_crs(nsw_lga)
proj4string <- crs_info$proj4string
epsg <- crs_info$epsg

@@

nsw_lga <- sf::st_transform(nsw_lga, crs = projstring)
```

```{r area_variable}
nsw_lga <- nsw_lga %>% mutate(logarea = log(AREA_SQKM))
```

```{r centroids}
## Use the create_centroids function to find polygon centroids:
centroids <- create_centroids(shp_sf = nsw_lga, sf_id = sf_id, projstring = projstring, verbose = FALSE)

centroids$logarea <- nsw_lga$logarea

@@

```{r, fig.cap="NSW Local Government Areas and derived centroids."}
fort_nsw <- fortify_sfc(nsw_lga, projstring = projstring)

end_cents <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude, colour = logarea), data= centroids) + theme_void() + coord_equal()
end_cents
```

@@

```{r bbox}
## Not necessary when using `create_hexmap`:
## create the bounding box from the set of polygon centroids:
bbox <- tibble::tibble(min = c(min(centroids$longitude), min(centroids$latitude)),
        max = c(max(centroids$longitude), max(centroids$latitude)))

@@

```{r grid}
grid <- tibble::as_tibble(expand.grid(hex_long = seq(bbox$min[1] - buffer_dist,
    bbox$max[1] + buffer_dist,
    hex_size),
    hex_lat = seq(bbox$min[2] - buffer_dist,
        bbox$max[2] + buffer_dist,
        hex_size)))
```

```{r plot_grid}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude), data= centroids, colour = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75) 
```

@@

```{r shift}
# Find every second latitude
shift_lat <- grid %>% dplyr::select(hex_lat) %>%
    dplyr::distinct() %>%
    dplyr::filter(dplyr::row_number() %% 2 == 1) %>% unlist()

@@

```{r plot_shift}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroids, fill = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75) 
```

@@

```{r grid_id}
full_grid <- grid %>%
    mutate(hex_long_int = dense_rank(hex_long)-1,
        hex_lat_int = dense_rank(hex_lat)-1)
```

@@

```{r buffer}
grid <- create_buffer(centroids, full_grid, hex_size, buffer_dist)
```

@@

### Step 3: Filtering the grid

Only the grid points between the rolling average of the minimum and maximum centroid values are kept, for each row and column of the grid.

```{r plot_buffer}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroids, fill = "#b2df8a", fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = full_grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75)
```

@@

```{r split_centroids}
# Split the centroid data set
centroids <- centroids %>% 
  nest(longitude, latitude) %>% 
  mutate(closest = map(data, closest_focal_point, focal_points = focal_points)) %>% 
  unnest(data, closest) %>% 
  arrange(focal_distance)
```

@@

```{r f_dist}
## Not necessary when using `create_hexmap`:
## Consider hexagon points up to a certain distance from centroid

@@

```{r, eval = FALSE}
hexmap_allocation <- allocate(centroids = centroids,
        hex_grid = grid,
        hex_size = hex_size,
        filter_dist = filter_dist,
        width = 15,
        focal_points = focal_points,
        verbose = TRUE,
        id = sf_id)
```

@@

```{r arrange}
s_centroids <- centroids %>% arrange(focal_distance)

@@

```{r orig_dist}
# keep value to reset expanded distances
expand_dist <- filter_dist
```

@@

Keep only the available hexagon points, this will prevent multiple areas being allocated to the same hexagon.

```{r available_points}
# filter for only the available hex grid points
hex_grid <- grid %>% filter(!assigned)
```

### Step 2: Filter the grid points for those closest to the centroid

This will allow only the closest points, that are not assigned, to be considered.

```{r close_points}
# filter grid for avaiable points
centroid1 <- centroids %>% head(1)

@@

Create a box filter

```{r, echo = FALSE}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroid1, fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)

@@

Create circle filter

```{r}
    hex_grid <- hex_grid %>%
        rowwise %>%
        mutate(
        hex_lat_c = hex_lat - flat,
        hex_long_c = hex_long - flong) %>%
        mutate(hyp = ((hex_lat_c^2) + (hex_long_c^2))^(1/2))

@@

Filter for angle within circle, relates hexagon location to focal point, Sydney.

@@

The allocation is returned and combined with the data relating to each polygon.

```{r}
hex_points_df <- centroid_allocation %>% 
  group_by(LGA_CODE11) %>%
  nest() %>%
  mutate(hex_points = 
      map(.x = data,.f = function(a) { 
        fortify_hexagon(data = a, hex_size = hex_size)})) %>%
  unnest(data) %>%
  unnest(hex_points)
```

@@

```{r}
end_hex <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") +
  theme_void() + coord_equal() +
    geom_hex(aes(x = hex_long, y = hex_lat, fill = AREA_SQKM), data = hexmap_lga, 
      position = "identity", stat = "identity")

gridExtra::grid.arrange(end_hex, end_cents)
```

@@

Examples of Melbourne, nsw, Australia
At different levels, distance from original centroids, change in area

LGA of all Australia
SA2 for nsw
